This is an example of article title

### 2019-05-31 paper/paper.tex
most of the results and nomenclature of ChipSort our favorite algorithm

\subsection{Vectorized Combsort}
%
One of the most peculiar aspects of ChipSort, taken directly from AA-sort~\cite{DBLP:conf/IEEEpact/InoueMKN07,DBLP:journals/pvldb/InoueT15}, is the use of the Combsort algorithm. Combsort is essentially a generalization of Bubble sort in the same way that Shell sort generalizes insertion sort~\cite{dobosiewicz1980efficient,Lacey:1991:FES:117187.117218,DBLP:books/lib/Knuth98a,DBLP:books/daglib/0023376}. The array is sweeped multiple times, and pairs of values at a given distance from each other are compared and stored in order. The distance is reduced at each pass by a factor of $\frac{4}{3}$. This algorithm lends itself easily to vectorization: performing the comparisons over vectors of size $V$ effectively allows us to run the algorithm over $V$ lists simultaneouly.

In practice, ChipSort performs a number of other operations before and after using Combsort. The {\tt chipsort\_medium!} function consists of the following steps:

\begin{enumerate}
\item Apply a sorting network on $K$ blocks of $J$ vectors of size $V$.
\item Vectorized Combsort until the interval size is 1.
\item Transpose blocks.
\item Transpose data in-place from $K$ blocks of $J$ vectors into $J$ blocks of $K$ vectors.
\item Vectorized Combsort again.
\item Sort blocks again.
\item Insertion sort.
\end{enumerate}

After the first two steps the result is essentially a matrix where the $K\times J$ columns are the vectors, and each row contains an approximately sorted sequence. The transpose steps reshape each row into a new block. Each row is processed independently at first, and the transpose allows elements from distinct groups to interact.

The last step is necessary to ensure the whole array is sorted, not vieved anymore as a matrix with independent rows. Insertion sort also presents a better performance than sticking to Combsort until the end, what would essentially mean using bubble sort once the interval becomes 1. Because the array is now approximately sorted, it is expected at this time that insertion sort does not have to move any out-of-order elements very far from where they are found.

@@

\section{Experiments}
\label{sec:experiments}
%
\begin{figure}[htb]
\centerline{\includegraphics[width=0.99\linewidth]{fig/chipsort-bench-8k.pdf}}
\centerline{\includegraphics[width=0.99\linewidth]{fig/chipsort-bench-256k.pdf}}
\caption{Empirical CDF from {\tt chipsort\_medium!} with 8k {\tt UInt32} values and {\tt chipsort\_large} with 256k values. Results from 10,000 measurements. Speedups are 14\% and 3\%.}
\label{fig:bench-cdf}
\end{figure}

@@

>>>>

\subsection{Vectorized Combsort}
%
One of the most peculiar aspects of ChipSort, taken directly from AA-sort~\cite{DBLP:conf/IEEEpact/InoueMKN07,DBLP:journals/pvldb/InoueT15}, is the use of the Combsort algorithm. Combsort is essentially a generalization of Bubble sort in the same way that Shell sort generalizes Insertion sort~\cite{dobosiewicz1980efficient,Lacey:1991:FES:117187.117218,DBLP:books/lib/Knuth98a,DBLP:books/daglib/0023376}. The array is swept multiple times, and pairs of values at a given distance from each other are compared and stored in order. The distance is reduced at each pass by a factor of $\frac{4}{3}$. This algorithm lends itself easily to vectorization: performing the comparisons over vectors of size $V$ effectively allows us to run the algorithm over $V$ lists simultaneously.

ChipSort contains two functions that utilize Combsort. The first is {\em chipsort!}, a serial implementation of the algorithm which is vectorized implicitly by the compiler optimizations. {\tt chipsort!} starts like Combsort until the interval size is 1, when it switches to Insertion sort. This is similar to what is often done in practical implementations of Quicksort, as in the Julia standard library, switching to Insertion sort when the array becomes too small. The difference is that Combsort is not a divide-and-conquer approach, and therefore Insertion sort is applied only once through the whole array.

The other function is {\tt chipsort\_medium!}, an explicitly vectorized implementation utilizing {\tt SIMD.jl}. This second function also performs a number of other operations before and after using Combsort. It consists of the following steps:
\begin{enumerate}
\item Apply a sorting network on $K$ blocks of $J$ vectors of size $V$.
\item Vectorized Combsort until the interval size is 1.
\item Transpose blocks.
\item Transpose data in-place from $K$ blocks of $J$ vectors into $J$ blocks of $K$ vectors.
\item Vectorized Combsort again.
\item Sort blocks again.
\item Insertion sort.
\end{enumerate}

After the first two steps the result is essentially a matrix where the $K\times J$ columns are the vectors, and each row contains an approximately sorted sequence. The transpose steps reshape each row into a new block. Each row is processed independently at first, and the transpose allows elements from distinct groups to interact. The last step is necessary to ensure the whole array is sorted, not viewed anymore as a matrix with independent rows.

In both implementations switching to Insertion sort presents a better performance than sticking to Combsort until the end. It essentially means using Bubble sort once the interval becomes 1, therefore looking for an alternative seems like a good idea. Insertion sort should fare well because the array is now approximately sorted, and it should not have to move any out-of-order elements very far from where they are found. This utilization of Insertion sort as a final stage to Combsort does not seem to have been considered in the literature before, and may be the main contribution of the present work.

@@

\section{Experiments}
\label{sec:experiments}
%
This section reports experiments with the methods offered by the ChipSort library, as well as alternative baseline methods. The term {\it method} can be understood in this article as both the sorting algorithm and their actual implementations in the Julia language. The specific methods tested were:
\begin{itemize}
\item {\tt chipsort\_small!} in-register, based on sorting networks.
\item {\tt chipsort\_medium!} vectorized Combsort, plus additional steps, finishing with Insertion sort.
\item {\tt chipsort\_large} merge-sort with a multi-way merge tree, starting from chunks processed by {\tt chipsort\_medium!}.
\item {\tt chipsort\_serial!} non-vectorized, serial Combsort finishing with Insertion sort when the interval becomes 1.
\item {\tt insertion\_sort!} Insertion sort taken from the Julia library.
\item {\tt sort!} the Julia standard sort, Quicksort finishing with Insertion sort for small arrays (20 elements).
\end{itemize}

The function names with a {\tt !} imply in-place operation. The output array allocation time for {\tt chipsort\_large} was not discounted in the experiments. While it may it not be irrelevant, it is small enough to be ignored with no effect to our conclusions.

\subsection{Small arrays}
In our first experiment we evaluate the sorting of small arrays, with only 64 32-bit values. The {\tt chipsort\_small} method based on sorting and bitonic merge networks was compared to serial Insertion sort, the Julia sort and {\tt chipsort!}. The small running time makes it challenging to perform proper measurements in this experiment. Our task was therefore to sort 128 sequences of 64 contiguous values in memory, forming a single block of $2^{13}$ random entries. The in-place sorting functions were applied in sequence, in a for-loop. The results are in Table~\ref{tab:bench-small}.

The {\tt chipsort\_small} function achieved a significant speed-up of over 3 times compared to Insertion sort or the Julia standard function, and twice over {\tt chipsort!}.

\begin{table}[h]
\tabcolsep22pt
\tbl{Sorting $128 \times 64$ {\tt UInt32} elements.}{
\begin{tabular}{|l|r|}\hline
Method & Median time\\\hline
ChipSort small & 57.206 $\mu$s\\\hline
ChipSort & 133.510 $\mu$s\\\hline
Insertion sort & 194.679 $\mu$s\\\hline
Julia standard & 222.910 $\mu$s\\\hline
\end{tabular}}
\label{tab:bench-small}
\end{table}

This first experiment already demonstrates the potential of {\tt chipsort!}, which presented a speedup of at least 46\% compared to the other alternatives, even though the function based on sorting networks was clearly faster.

\subsection{Medium and large arrays}
In our second experiment we compared many different methods on inputs of exponentially increasing sizes, from $2^6$ up to $2^{20}$. Figure~\ref{fig:bench-time} displays the resulting measurements. Although the curves look too concentrated in this visualization, it is possible to notice that {\tt chipsort!} consistently outperforms the Julia standard library {\tt sort!} by a factor of 80\% to 100\%, or up to twice as fast. We can also observe that Insertion sort soon becomes too inefficient.

\begin{figure}[htb]
\centerline{\includegraphics[width=0.99\linewidth]{fig/chipsort-bench-time.pdf}}
\caption{Running time of the different sorting methods studied.}
\label{fig:bench-time}
\end{figure}

Figure~\ref{fig:bench-curves} shows a different visualization from the same data in Figure~\ref{fig:bench-time}, with the times divided by the input size at each point. Here we can see more clearly how {\tt chipsort\_medium} and {\tt chipsort\_large} slightly outperform {\tt chipsort!} in a few cases.

@@

Figure~\ref{fig:bench-cdf} displays the measured times for

 Speedups are 7.3\% and 4.4\%.

\begin{figure}[htb]
\centerline{\includegraphics[width=0.99\linewidth]{fig/chipsort-bench-8k.pdf}}
\centerline{\includegraphics[width=0.99\linewidth]{fig/chipsort-bench-256k.pdf}}
\caption{Empirical CDF from different methods with 8k and 256k {\tt UInt32} values. Results from 10,000 measurements.}
\label{fig:bench-cdf}
\end{figure}
